---
title: "p8105_mtp_yz4184"
author: "Yunlin Zhou"
date: "10/28/2021"
output: html_document
---
```{r setup, include = FALSE, echo = FALSE}
library(tidyverse)
library(readxl)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

```{r}
# Import data

horns_df_ori = read_excel ("./p8105_mtp_data.xlsx", range = "A9:I1230")

```


```{r}
# Clean the data

horns_df  = janitor::clean_names(horns_df_ori)

horns_df = horns_df %>%
  mutate(horns_df,
    sex = recode(sex, `0` = "female", `1` = "male"),
    sex = factor(sex, levels = c("female", "male")),
    age_group = recode(age_group, 
                       `1` = "18-",
                       `2` = "18-30",
                       `3` = "31-40",
                       `4` = "41-50",
                       `5` = "51-60",
                       `6` = "60+",
                       `7` = "60+",
                       `8` = "60+"),
      age_group = factor(age_group,levels = 
                              c("18-30",
                                "31-40",
                                "41-50",
                                "51-60",
                                "60+")),
    eop_size = recode(eop_size, 
                       `0` = "0-5",
                       `1` = "5-10",
                       `2` = "10-15",
                       `3` = "15-20",
                       `4` = "20-25",
                       `5` = "25+"),
    eop_size =factor(eop_size,
                           levels = 
                              c("0-5",
                                "5-10",
                                "10-15",
                                "15-20",
                                "20-25",
                                "25+")),
    eop_visibility_classification = recode(eop_visibility_classification, 
                       `0` = "0",
                       `1` = "0-5",
                       `2` = "5+"),
    eop_visibility_classification =
      factor(eop_visibility_classification, levels = c("0", "0-5", "5+")),
    fhp_category = recode(fhp_category,
                          `0` = "0-10",
                       `1` = "10-20",
                       `2` = "20-30",
                       `3` = "30-40",
                       `4` = "40-50"),
    fhp_category = factor(fhp_category, levels = c("0-10", "10-20", "20-30", "30-40", "40-50")),
    eop_size_mm = replace (eop_size_mm, 
                           is.na (eop_size_mm), 0),
    eop_shape = replace (eop_shape, 
                           is.na (eop_shape), 0))%>%
  relocate(sex, age_group)%>%
  arrange(sex,age)%>%
  filter(age_group != "18-")

```


According to the original table, I replaced the group number with accurate range, and fill in missing values with data. Also I tidied up column names, relocated the column orders and reorder the rows by sex and age. 

Key variables:age, fhp_size_mm, eop_size

```{r results='hide', echo = FALSE}
# Describe the resulting dataset

ncol(horns_df)

nrow(horns_df)

sex_num = count(horns_df,sex)
knitr::kable(sex_num)

age_num = count(horns_df, age_group)
knitr::kable(age_num)

```

There are `r ncol(horns_df) ` variables and `r nrow(horns_df) ` participants which included 614 female and 607 male. There are 2 people under 18, 303 people from 18 to 30, 204 people from 31 to 40, 207 people from 41 to 50, 200 people from 51 to 60 and 305 people above 60. The key vairaibles are sex, FHP and age.



```{r}
# Age and gender distribution

age_gender = horns_df%>%
  group_by(sex, age_group)%>%
  count()

knitr::kable(age_gender)
```


```{r}
# issues in the available data

# In FHP category, there is a number not belonging to any group.

strange_fhp = horns_df_ori%>%
  janitor::clean_names()%>%
  mutate(fhp_category = as.numeric(fhp_category))%>%
  filter(fhp_category > 7)
knitr::kable(strange_fhp)

# age not mapping to age_group
horns_df_ori%>%
  janitor::clean_names()%>%
ggplot(aes(x = age_group, y = age))+geom_point()

# In EOP Size, there is a number not belonging to any group.

strange_eop = horns_df_ori%>%
  janitor::clean_names()%>%
  mutate(eop_size = as.numeric(eop_size))%>%
  filter(eop_size > 5)
knitr::kable(strange_eop)

```


# Problem 2 

```{r}

library(patchwork)
# FHP histogram
fhp_his_plot = 
  horns_df %>%
  ggplot(aes(x = fhp_size_mm, fill = sex))+
  geom_histogram(position = "dodge", binwidth = 2)+ 
  labs(
    title = "FHP size histogram",
    x = "FHP size (mm)",
    y = "Different FHP size number"
  )



# The rate of enlarged EOP in each age group

rate_df = horns_df %>%
  count(age_group, sex, eop_size)

eop = rate_df%>%
  filter(eop_size%in%c("10-15","15-20","20-25","25+"))%>%
  group_by(age_group,sex)%>%
  summarise(eop_num = sum(n))

total = rate_df%>%
  group_by(age_group,sex)%>%
  summarise(total_num = sum(n))

rate_df = left_join(eop, total)%>%
  mutate(rate = eop_num/total_num,
         age_group_num = recode(age_group,
                                `18-30` = 1,
                                `31-40` = 2,
                                `41-50` = 3,
                                `51-60` = 4,
                                `60+` = 5))

rate_df_plot = ggplot(rate_df, aes(x = age_group_num, y = rate, color =sex))+
  geom_point()+
  geom_line()+
  labs(
    title = "The rate of enlarged EOP in each age group",
    x = "Age group (years)",
    y = "The rate of enlarged EOP"
  ) + 
  scale_x_continuous(
    labels = c("18-30", "31-40", "41-50", "51-60", "60+")
  ) 

fhp_his_plot + rate_df_plot

```

As  shown in the left plot, the distributions are both skewed right and the data seem to be centered around 25. Male are more likely to have higher FHP size.

The rate of EEOP is different from the plot in the article. Younger and older people have higher EEOP rate than others.

```{r fig.asp = 2}
# The association between FHP size and EOP size in each age and sex group

eop_fhp = 
horns_df%>%
  filter(age_group != "18-")%>%
ggplot(aes(x = eop_size_mm, y = fhp_size_mm))+
  geom_point()+
  labs(
    title = "The association between FHP size and EOP size",
    x = "EOP size (mm)",
    y = "FHP size (mm)"
  ) + 
  scale_x_continuous(
    breaks = c(10, 20, 30),
    labels = c("10", "20", "30")
  )+geom_smooth()+
  facet_wrap(vars(sex,age_group), ncol = 2)

eop_fhp
```

According to the plot, there is no obvious linear association between EOP size and FHP size. Older people and male are more likely to have higher EOP size and FHP size. 

# Problem 3

```{r}
# Sample sizes in each age group

sample_size_each = horns_df%>%count(age_group)
knitr::kable(sample_size_each)

sample_size = length(horns_df$sex)
print(sample_size)

sample_range = range(horns_df$age)
print(sample_range)

```

The sample size and range are approximately the same to that mentioned in article.

```{r}
# Mean and standard deviations for FHP size

# mean and sd of male
horns_df%>%
  filter(sex == "male")%>%
  select(fhp_size_mm)%>%
t.test(alternative = 'two.sided')

sd_male_df = horns_df%>%
  filter(sex == "male")%>%
  mutate(fhp_size_mm = replace(fhp_size_mm, is.na(fhp_size_mm),0))%>%
  select(fhp_size_mm)

sd_male = as.numeric(unlist(sd_male_df))%>%
  sd()%>%
  print()



# mean and sd of female

horns_df%>%
  filter(sex == "female")%>%
  select(fhp_size_mm)%>%
t.test(alternative = 'two.sided')

sd_female_df = horns_df%>%
  filter(sex == "female")%>%
  mutate(fhp_size_mm = replace(fhp_size_mm, is.na(fhp_size_mm),0))%>%
  select(fhp_size_mm)

sd_female = as.numeric(unlist(sd_female_df))%>%
  sd()%>%
  print()

```

The mean and sd in article are close to those calculated by my data.


EEOP definition: EOP (external occipital protuberance) which exceeded 10 mm in size

eop_size and eop_size_mm could be used to evaluate the claim.

```{r}
# the prevalence of EEOP

eeop_num = horns_df%>%
  filter(eop_size %in% c("10-15", "15-20", "20-25", "25+"))%>%
  nrow()%>%
  print()

eeop_prev = (eeop_num/nrow(horns_df))%>%print()

```



```{r}
# FHP in older subjects

# The FHP trend

fhp_density = ggplot(horns_df, aes(x = fhp_size_mm, color = age_group))+
  geom_density()+
  labs(
    title = "Density of FHP size in different age group",
    x = "FHP size",
    y = "Density"
  )

fhp_density

```

As shown in the plot, in 60+ age group, the density is higher in large FHP size.

```{r}

# FHP >40 mm rate in older subjects

FHP_older = horns_df%>%
  filter(age_group == "60+")%>%
  nrow()

FHP_high_older = horns_df%>%
  filter(fhp_size_mm > 40,
         age_group == "60+")%>%nrow()

fhp_older_rate = (FHP_high_older/FHP_older)%>%print()

```
The rate I calculated is close to 34.5%.

# Problem 4

